Bayesian Nonparametric System Reliability\\ using Sets of Priors

### 2016-07-04 ijar-npb-ms.tex
some minor tweaks in Louis' additions

If $\frac{s}{N + m - 1} < y < \frac{s + m - 1}{N + m - 1}$ then 
Theorem 2 cannot determine stochastic dominance.  The following Lemma which is slightly
more computationally costly, but still much faster than an exhaustive search, may be able to determine first-order
stochastic dominance in such situations.
\begin{lemma}
  \label{lem:n}
  Let $\beta_n$ denote the Beta-Binomial distribution as in Theorem \ref{thm:n}.  Define
  \[ \mathcal{L}_{\nu,\nl}(l) := \frac{p(l \mid y, \nu, m, s, N)}{p(l \mid y, \nl, m, s, N)} \]
  Then,
  \[ \left. \begin{aligned}
       \mathcal{L}_{\nu,\nl}(0) \le 1 \\
       \mathcal{L}_{\nu,\nl}(m) \ge 1 \\
    \end{aligned} \right\}
    \implies \beta_{\nu} \ge_{\mathrm{st}} \beta_{\nl} \]
  and
  \[ \left. \begin{aligned}
       \mathcal{L}_{\nu,\nl}(0) \ge 1 \\
       \mathcal{L}_{\nu,\nl}(m) \le 1 \\
    \end{aligned} \right\}
    \implies \beta_{\nu} \le_{\mathrm{st}} \beta_{\nl} \]
\end{lemma}
The proof is provided in Appendix \ref{ap:proofs}, p.\pageref{prf:n2}.
In the cases where neither Theorem \ref{thm:n} or Lemma \ref{lem:n} apply,
the entire posterior system reliability function must be optimised to 
find the minima/maxima.
In practice, in the examples to be presented in the sequel, Theorem 
\ref{thm:n} and Lemma \ref{lem:n} do provide guarantees of first-order
stochastic dominance for the vast majority of time points, $t$,
substantially lowering the computational costs of performing the
minimisation/maximisation involved in finding the sets of system 
reliability functions compared to either numerical optimisation or
an exhaustive grid search (which would get exponentially slower in 
the number of different components).

Appendix \ref{ap:needtoopt} provides a detailed discussion of exactly how often these bounds hold and how often one must resort to numerical optimisation.

Thus,
%For $\lRsys(\cdot)$, the minimum is achieved at
%$\nktzu$ if $\yktzl(n_k+m_k-1)-s^k_t-m_k < 0$ and is achieved at 
%$\nktzl$ if $\yktzl(n_k+m_k-1)-s^k_t-m_k > 0$.
%Conversely, for $\uRsys(\cdot)$, the maximum is achieved at 
%$\nktzu$ if $\yktzu(n_k+m_k-1)-s^k_t-m_k > 0$ and is achieved at 
%$\nktzl$ if $\yktzu(n_k+m_k-1)-s^k_t-m_k < 0$.
\begin{align}
\lefteqn{\lRsys(t \mid \vec{t}^1, \ldots, \vec{t}^K)} \nonumber \\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \nonumber \\ & \hspace*{8ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nktz\yktz + s^k_t, m_k - l_k + \nktz(1-\yktz) + n_k - s^k_t)}{B(\nktz\yktz + s^k_t, \nktz(1-\yktz) + n_k - s^k_t)}
    \nonumber\\
 &= \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \nonumber \\ & \hspace*{8ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nktzo\yktzl + s^k_t, m_k - l_k + \nktzo(1-\yktzl) + n_k - s^k_t)}{B(\nktzo\yktzl + s^k_t, \nktzo(1-\yktzl) + n_k - s^k_t)} \label{eq:LwrSysPost}\\
 &\mathrm{where} \nonumber \\
 &\nktzo = \left\{ \begin{aligned}
   \nktzu & \quad\mbox{if}\ \yktzl < \frac{s^k_t}{n_k + m_k - 1} \vee \Big( \mathcal{L}_{\nktzu,\nktzl}(0) \ge 1 \wedge \mathcal{L}_{\nktzu,\nktzl}(m) \le 1 \Big) \\
   \nktzl & \quad\mbox{if}\ \yktzl > \frac{s^k_t + m_k - 1}{n_k + m_k - 1} \vee \Big( \mathcal{L}_{\nktzu,\nktzl}(0) \le 1 \wedge \mathcal{L}_{\nktzu,\nktzl}(m) \ge 1 \Big) \\
   & \quad\mbox{optimised otherwise}
 \end{aligned} \right. \nonumber
\end{align}

@@

We also consider a simplified automotive brake system.
The master brake cylinder (M) activates all four wheel brake cylinders (C1 -- C4),
which in turn actuate a braking pad assembly each (P1 -- P4).
The hand brake mechanism (H) goes directly to the brake pad assemblies P3 and P4;
the car brakes when at least one brake pad assembly is actuated.
All values for $\Phi \not\in \{0,1\}$ are given in Table~\ref{tab:brake-survsign}.
\begin{table}
\centering
\begin{tabular}{cccclcccccl}
  \toprule
M & H & C & P & $\Phi$ & \quad & M & H & C & P & $\Phi$\\ 
  \midrule
1 & 0 & 1 & 1 & 0.25 & & 1 & 0 & 2 & 1 & 0.50 \\ 
1 & 0 & 1 & 2 & 0.50 & & 1 & 0 & 2 & 2 & 0.83 \\ 
1 & 0 & 1 & 3 & 0.75 & & 1 & 0 & 3 & 1 & 0.75 \\ 
0 & 1 & 0 & 1 & 0.50 & & 1 & 1 & 0 & 1 & 0.50 \\ 
0 & 1 & 0 & 2 & 0.83 & & 1 & 1 & 0 & 2 & 0.83 \\ 
0 & 1 & 1 & 1 & 0.62 & & 1 & 1 & 1 & 1 & 0.62 \\ 
0 & 1 & 1 & 2 & 0.92 & & 1 & 1 & 1 & 2 & 0.92 \\ 
0 & 1 & 2 & 1 & 0.75 & & 1 & 1 & 2 & 1 & 0.75 \\ 
0 & 1 & 2 & 2 & 0.97 & & 1 & 1 & 2 & 2 & 0.97 \\ 
0 & 1 & 3 & 1 & 0.88 & & 1 & 1 & 3 & 1 & 0.88 \\ 
   \bottomrule
\end{tabular}
\caption{Survival signature values $\not\in \{0,1\}$ for the simplified automotive brake system depicted in Figure~\ref{fig:brakesystem}.
}
\label{tab:brake-survsign}
\end{table}
The system layout is depicted in Figure~\ref{fig:brakesystem},
together with prior and posterior sets of reliability functions for the four component types and the complete system.
Observed lifetimes from test data are indicated by tick marks in each of the four component type panels,
where $n_\text{M}=5$, $n_\text{H}=10$, $n_\text{C}=15$, and $n_\text{P}=20$.
We assume $[\ntzl{\text{M}},\ntzu{\text{M}}] = [1,8]\ \forall t$,
and $[\nktzl, \nktzu] = [1,2]$ for $k \in \{\text{H, C, P}\}$ and all $t$.
Prior functioning probability bounds for M are based on
a Weibull cdf with shape $2.5$ and scales $6$ and $8$ for the lower and upper bound, respectively.
The prior bounds for P can be seen as the least committal bounds
derived from an expert statement of $\ytz{\text{P}} \in [0.5, 0.65]$ for $t=5$ only.
%
For H, near-noninformative prior functioning probability bounds have been selected;
with the upper bound for P being approximately one for $t \le 5$ as well,
the prior upper system reliability bound for $t \le 5$ is close to one, too,
since the system can function on H and one of P1 -- P4 alone.
Note that the posterior functioning probability interval for M
is wide not only due to the limited number of observations,
but also because $\ntzu{\text{M}} = 8$ and the prior-data conflict reaction.

@@

In this example, the theory presented in Section \ref{sec:setsofrel} enabled avoiding numerical optimisation in 91.9\% of cases for $\lRsys(\cdot)$ and in 95.7\% of cases for $\uRsys(\cdot)$.  Further detail is in Appendix \ref{ap:needtoopt}.

@@

Given some unknown failure probability $p$, which gives rise to the observed data $s$, the probability that Theorem 1 does not apply is
\begin{align*}
  &P(y(N+m-1)-m+1 < s < y(N+m-1)) \\
  &\qquad \qquad \qquad \qquad = \sum_{i = \lceil y(N+m-1)-m+1 \rceil}^{\lfloor y(N+m-1) \rfloor} \binom{N}{i} p^i (1-p)^{N-i}
\end{align*}
Likewise the probability that Lemma 3 does not apply is
\begin{align*}
  &P(\mathcal{L}_{\nu,\nl}(0) \lesseqgtr 1 \cap \mathcal{L}_{\nu,\nl}(m) \gtreqless 1) = \sum_{s \in \mathcal{S}} \binom{N}{s} p^s (1-p)^{N-s} \\
  &\qquad\mbox{where}\quad \mathcal{S} = \left\{ s : \frac{p(0 \mid y, \nu, m, s, N)}{p(0 \mid y, \nl, m, s, N)} \lesseqgtr 1 \cap \frac{p(m \mid y, \nu, m, s, N)}{p(m \mid y, \nl, m, s, N)} \gtreqless 1 \right\}
\end{align*}

@@

\begin{figure}
\includegraphics[width=\textwidth]{need-to-opt-3}
\caption{Plots of the probability that numerical optimisation is required when $y=0.4$, $N=100$, $\nl=1$ and $\nu=5$ for differing numbers of component in the system, $m$, as a function of the unknown failure probability, $p$.}
\label{fig:need-to-opt-3}
\end{figure}

@@

The worst case scenario is that the prior, $y$, is exactly equal to the true unknown probability of failure, $p$.  Regardless of $y$, optimisation is required less frequently with larger test dataset sizes (increasing $N$), but offset by increasing numbers of the same component in a system (increasing $m$).  Not shown is that $\nl$ and $\nu$ clearly do not affect the probability that Theorem 2 does not apply, but also the impact on Lemma 3 is negligible.

@@

Note that all these analyses are for evaluation of a single fixed time point at which $\lRsys(\cdot)$ or $\uRsys(\cdot)$ computed.  Of relevance to the overall compute time is the behaviour over time, so that is seems unlikely the worst case scenario will be occur for long runs of the times under analysis.

@@

Figure \ref{fig:need-to-opt-8} shows the changing worst case probability (subject to $|y-p|>0.1$) that optimisation will be required over time for the actual choice of prior and the values of $m, N, \nu$ and $\nl$ in the brake example from the main body of the paper.  Note in particular that when $m=1$, Lemma 3 is always applies and optimisation is never required.  This is true for all values of $N$ and $y$.

>>>>

If $\frac{s}{N + m - 1} < y < \frac{s + m - 1}{N + m - 1}$ then 
Theorem 2 cannot determine stochastic dominance.  The following Lemma which is slightly
more computationally costly, but still much faster than an exhaustive search, may be able to determine first-order
stochastic dominance in such situations.
\begin{lemma}
  \label{lem:n}
  Let $\beta_n$ denote the Beta-Binomial distribution as in Theorem \ref{thm:n}.  Define
  \[ \mathcal{L}_{\nu,\nl}(l) := \frac{p(l \mid y, \nu, m, s, N)}{p(l \mid y, \nl, m, s, N)} \]
  Then,
  \[ \left. \begin{aligned}
       \mathcal{L}_{\nu,\nl}(0) \le 1 \\
       \mathcal{L}_{\nu,\nl}(m) \ge 1 \\
    \end{aligned} \right\}
    \implies \beta_{\nu} \ge_{\mathrm{st}} \beta_{\nl} \]
  and
  \[ \left. \begin{aligned}
       \mathcal{L}_{\nu,\nl}(0) \ge 1 \\
       \mathcal{L}_{\nu,\nl}(m) \le 1 \\
    \end{aligned} \right\}
    \implies \beta_{\nu} \le_{\mathrm{st}} \beta_{\nl} \]
\end{lemma}
The proof is provided in Appendix \ref{ap:proofs}, p.\pageref{prf:n2}.
In the cases where neither Theorem \ref{thm:n} or Lemma \ref{lem:n} apply,
the entire posterior system reliability function must be optimised to 
find the minima/maxima.
In practice, in the examples to be presented in the sequel, Theorem 
\ref{thm:n} and Lemma \ref{lem:n} do provide guarantees of first-order
stochastic dominance for the vast majority of time points, $t$,
substantially lowering the computational costs of performing the
minimisation/maximisation involved in finding the sets of system 
reliability functions compared to either numerical optimisation or
an exhaustive grid search (which would get exponentially slower in 
the number of different components).
%
Appendix \ref{ap:needtoopt} provides a detailed discussion of exactly how often these bounds hold and how often one must resort to numerical optimisation.

Thus, due to Theorems~\ref{thm:y}, \ref{thm:n} and Lemma~\ref{lem:n}, \eqref{eq:LwrSysPostA} transforms to
%For $\lRsys(\cdot)$, the minimum is achieved at
%$\nktzu$ if $\yktzl(n_k+m_k-1)-s^k_t-m_k < 0$ and is achieved at 
%$\nktzl$ if $\yktzl(n_k+m_k-1)-s^k_t-m_k > 0$.
%Conversely, for $\uRsys(\cdot)$, the maximum is achieved at 
%$\nktzu$ if $\yktzu(n_k+m_k-1)-s^k_t-m_k > 0$ and is achieved at 
%$\nktzl$ if $\yktzu(n_k+m_k-1)-s^k_t-m_k < 0$.
\begin{align}
\lefteqn{\lRsys(t \mid \vec{t}^1, \ldots, \vec{t}^K)} \nonumber \\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \nonumber \\ & \hspace*{8ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nktz\yktz + s^k_t, m_k - l_k + \nktz(1-\yktz) + n_k - s^k_t)}{B(\nktz\yktz + s^k_t, \nktz(1-\yktz) + n_k - s^k_t)}
    \nonumber\\
 &= \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \nonumber \\ & \hspace*{8ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nktzo\yktzl + s^k_t, m_k - l_k + \nktzo(1-\yktzl) + n_k - s^k_t)}{B(\nktzo\yktzl + s^k_t, \nktzo(1-\yktzl) + n_k - s^k_t)} \label{eq:LwrSysPost}\\
 &\mathrm{where} \nonumber \\
 &\nktzo = \left\{ \begin{aligned}
   \nktzu & \quad\mbox{if}\ \yktzl < \frac{s^k_t}{n_k + m_k - 1} \vee \Big( \mathcal{L}_{\nktzu,\nktzl}(0) \ge 1 \wedge \mathcal{L}_{\nktzu,\nktzl}(m) \le 1 \Big) \\
   \nktzl & \quad\mbox{if}\ \yktzl > \frac{s^k_t + m_k - 1}{n_k + m_k - 1} \vee \Big( \mathcal{L}_{\nktzu,\nktzl}(0) \le 1 \wedge \mathcal{L}_{\nktzu,\nktzl}(m) \ge 1 \Big) \\
   & \quad\mbox{optimised otherwise}
 \end{aligned} \right. \nonumber
\end{align}

@@

We also consider a simplified automotive brake system.
The master brake cylinder (M) activates all four wheel brake cylinders (C1 -- C4),
which in turn actuate a braking pad assembly each (P1 -- P4).
The hand brake mechanism (H) goes directly to the brake pad assemblies P3 and P4;
the car brakes when at least one brake pad assembly is actuated.
All values for $\Phi \not\in \{0,1\}$ are given in Table~\ref{tab:brake-survsign}.
\begin{table}
\centering
\begin{tabular}{cccclcccccl}
  \toprule
M & H & C & P & $\Phi$ & \quad & M & H & C & P & $\Phi$\\ 
  \midrule
1 & 0 & 1 & 1 & 0.25 & & 1 & 0 & 2 & 1 & 0.50 \\ 
1 & 0 & 1 & 2 & 0.50 & & 1 & 0 & 2 & 2 & 0.83 \\ 
1 & 0 & 1 & 3 & 0.75 & & 1 & 0 & 3 & 1 & 0.75 \\ 
0 & 1 & 0 & 1 & 0.50 & & 1 & 1 & 0 & 1 & 0.50 \\ 
0 & 1 & 0 & 2 & 0.83 & & 1 & 1 & 0 & 2 & 0.83 \\ 
0 & 1 & 1 & 1 & 0.62 & & 1 & 1 & 1 & 1 & 0.62 \\ 
0 & 1 & 1 & 2 & 0.92 & & 1 & 1 & 1 & 2 & 0.92 \\ 
0 & 1 & 2 & 1 & 0.75 & & 1 & 1 & 2 & 1 & 0.75 \\ 
0 & 1 & 2 & 2 & 0.97 & & 1 & 1 & 2 & 2 & 0.97 \\ 
0 & 1 & 3 & 1 & 0.88 & & 1 & 1 & 3 & 1 & 0.88 \\ 
   \bottomrule
\end{tabular}
\caption{Survival signature values $\not\in \{0,1\}$ for the simplified automotive brake system depicted in Figure~\ref{fig:brakesystem}.
}
\label{tab:brake-survsign}
\end{table}
The system layout is depicted in Figure~\ref{fig:brakesystem},
together with prior and posterior sets of reliability functions for the four component types and the complete system.
Observed lifetimes from test data are indicated by tick marks in each of the four component type panels,
where $n_\text{M}=5$, $n_\text{H}=10$, $n_\text{C}=15$, and $n_\text{P}=20$.
We consider $301$ evenly spaced time points $t$ on $[0, 10]$,
assume $[\ntzl{\text{M}},\ntzu{\text{M}}] = [1,8]\ \forall t$,
and $[\nktzl, \nktzu] = [1,2]$ for $k \in \{\text{H, C, P}\}$ and all $t$.
Prior functioning probability bounds for M are based on
a Weibull cdf with shape $2.5$ and scales $6$ and $8$ for the lower and upper bound, respectively.
The prior bounds for P can be seen as the least committal bounds
derived from an expert statement of $\ytz{\text{P}} \in [0.5, 0.65]$ for $t=5$ only.
%
For H, near-noninformative prior functioning probability bounds have been selected;
with the upper bound for P being approximately one for $t \le 5$ as well,
the prior upper system reliability bound for $t \le 5$ is close to one, too,
since the system can function on H and one of P3 and P4 alone.
Note that the posterior functioning probability interval for M
is wide not only due to the limited number of observations,
but also because $\ntzu{\text{M}} = 8$ and the prior-data conflict reaction.

@@

In this example, the theory presented in Section \ref{sec:setsofrel} enabled avoiding numerical optimisation
for 91.9\% and 95.7\% of the $301$ considered time points for $\lRsys(\cdot)$ and $\uRsys(\cdot)$, respectively.
Further detail is in Appendix \ref{ap:needtoopt}.

@@

Given some unknown functioning probability $p$, which gives rise to the observed data $s$, the probability that Theorem 1 does not apply is
\begin{align*}
  &P(y(N+m-1)-m+1 < s < y(N+m-1)) \\
  &\qquad \qquad \qquad \qquad = \sum_{i = \lceil y(N+m-1)-m+1 \rceil}^{\lfloor y(N+m-1) \rfloor} \binom{N}{i} p^i (1-p)^{N-i}
\end{align*}
Likewise the probability that Lemma 3 does not apply is
\begin{align*}
  &P(\mathcal{L}_{\nu,\nl}(0) \lesseqgtr 1 \cap \mathcal{L}_{\nu,\nl}(m) \gtreqless 1) = \sum_{s \in \mathcal{S}} \binom{N}{s} p^s (1-p)^{N-s} \\
  &\qquad\mbox{where}\quad \mathcal{S} = \left\{ s : \frac{p(0 \mid y, \nu, m, s, N)}{p(0 \mid y, \nl, m, s, N)} \lesseqgtr 1 \cap \frac{p(m \mid y, \nu, m, s, N)}{p(m \mid y, \nl, m, s, N)} \gtreqless 1 \right\}
\end{align*}

@@

\begin{figure}
\includegraphics[width=\textwidth]{need-to-opt-3}
\caption{Plots of the probability that numerical optimisation is required when $y=0.4$, $N=100$, $\nl=1$ and $\nu=5$ for differing numbers of components in the system, $m$, as a function of the unknown failure probability, $p$.}
\label{fig:need-to-opt-3}
\end{figure}

@@

The worst case scenario is that the prior, $y$, is exactly equal to the true unknown probability of failure, $p$.  Regardless of $y$, optimisation is required less frequently with larger test dataset sizes (increasing $N$), but offset by increasing numbers of components of the same type in a system (increasing $m$).  Not shown is that $\nl$ and $\nu$ clearly do not affect the probability that Theorem 2 does not apply, but also the impact on Lemma 3 is negligible.

@@

Note that all these analyses are for evaluation of a single fixed time point at which $\lRsys(\cdot)$ or $\uRsys(\cdot)$ is computed.  Of relevance to the overall compute time is the behaviour over time, so that seems unlikely that the worst case scenario will occur for long runs of the times under analysis.

@@

Figure \ref{fig:need-to-opt-8} shows the changing worst case probability (subject to $|y-p|>0.1$) that optimisation will be required over time for the actual choice of prior and the values of $m, N, \nu$ and $\nl$ in the brake example from the main body of the paper.  Note in particular that when $m=1$, Lemma 3 *** always applies and optimisation is never required.  This is true for all values of $N$ and $y$.
